### 2020-04-20 survival.tex

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Survival Analysis
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Survival Analysis}
\label{section:survival}
The construction of a GPU life time is more complex than it initially
appears because the units are observed only at reboot time, because
most units were proactively replaced to prevent failure, and because
some units continue in operation after OTB and DBE events (a second
reboot may be successful).

@@

A key concept in survival analysis and in reliability analysis is
censoring, which is about using information from study subjects whose
exact failure times are not available or that have not failed. This
applies to our study because of proactive GPU replacement before
failure, because most units were still in operation when the system
was shut down, and also because of observation only at inspection
times (reboots). We use censoring concepts on the proactive
replacements and units still in operation at the end. This allows us
to use all of the GPU life time data, including units that did not
fail. But we ignore the inspection time censoring, treating removal
times as exact failure times to reduce the complexity of this
analysis. We expect that because of the volume of data, this would
hardly make a difference in our conclusions.

Kaplan-Meier survival analysis (KM) gives the probability of survival
beyond a given time. It is a nonparametric technique that makes no
specific failure model assumptions, such as Weibull, Exponential,
etc. It can also split the data into subpopulations and compute
separate survival curves.

If $T$ is taken to be the random variable of a GPU fail time, then its
probability density function is $f(t)$ and its cumulative distribution
function $F(t) = Pr\{T < t\}$ gives the probability that a GPU fails
by duration $t$. The survival function is then 
\begin{displaymath}
  S(t) = Pr\{T \geq t\} = 1 - F(t) = \int_t^\infty f(x)dx,
\end{displaymath}
which is the probability of being operational at duration
$t$.

We use reliability censoring concepts to use information from units
that were proactively removed before failing as well as those still in
operation till the end. We use the R package {\tt survival} for the KM
analysis.
\begin{figure*}
  \centering
  \includegraphics[width=7in]{figs/km_cage-node_a001.pdf}
  \caption{Comparison of the old and new batches, including survival
    differences based on {\tt cage} and {\tt node} GPU locations.}
  \label{fig:km-all-cage-node}
\end{figure*}

A sister concept to the survival function is the hazard function:
\begin{displaymath}
  H(t) = \frac{f(t)}{S(t)},
\end{displaymath}
which is what is sometimes called the "bathtub" curve in reliability.

The Cox proportional hazards (CPH) regression analys
\cite{Cox1972,Harrell2015} takes the hazard function above and a set
of covariates $x$ and models $k$s GPU hazard as 
\begin{displaymath}
  H_k(t) = H_0(t)e^{\sum\limits_{i=1}^n{\beta x}}
\end{displaymath}
That is, a base hazard rate multiplied by a function of
covariates. The key assumption in the CPH model is that the "hazards"
are multipliers on the "baseline hazard" but the estimate of the
baseline hazard is nonparametric, meaning it makes no specific
distributional assumption and just learns from the data (unlike
analyses that assume for example a Weibul or an exponential
model). The multiplicative assumption is that the hazard curves do not
cross and are multiples of each other. The multiplier is the *hazard
coefficient*. For example, if the baseline is *node0* and the *node2*
hazard is 2, then *node2* sees twice as many failures as *node0* on
average.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{figs/cox_o001.pdf}
  \caption{GPU hazard ratios from Cox regression model on {\tt old}
    batch.}
\end{figure}
\begin{figure}
  \includegraphics[width=\columnwidth]{figs/cox_n001.pdf}
  \caption{GPU hazard ratios from Cox regression model on {\tt new}
    batch.}
\end{figure}

>>>>

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Survival Analysis
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Survival Analysis}
\label{section:survival}
\renewcommand{\pkg}[1]{\textsf{#1}}
\begin{figure*}
  \centering
  \includegraphics[width=7in]{figs/km_cage-node_a001.pdf}
  \caption{Comparison of the old and new batches, including survival
    differences based on {\tt cage} and {\tt node} GPU locations.}
  \label{fig:km-all-cage-node}
\end{figure*}
The construction of a GPU life time is more complex than it initially
appears because the units are observed only at reboot time, because
most units were proactively replaced to prevent failure, and because
some units continue in operation after OTB and DBE events (a second
reboot may be successful).

@@

A key concept in survival analysis and in reliability analysis is
censoring, which is about using information from study subjects whose
exact failure times are not available or that have not failed. This
applies to our study because of proactive GPU replacement before
failure, because most units were still in operation when the system
was shut down, and also because of observation only at inspection
times (reboots). We use censoring concepts on the proactive
replacements and units still in operation at the end. This allows us
to use all of the GPU life time data, including units that did not
fail. But we ignore the inspection time censoring, treating removal
times as exact failure times to reduce the complexity of this
analysis. We expect that because of the volume of data, this would
make much difference in our conclusions. However, we are making our
data publicly available and expect that others, especially in the
survival analysis community, will dive deeper.

Kaplan-Meier survival analysis (KM) gives the probability of survival
beyond a given time. It is a nonparametric technique that makes no
specific failure model assumptions, such as Weibull, Exponential,
etc. The technique is able to use censored observations and can also
split the data into subpopulations and compute separate survival
curves.

If $T$ is taken to be the random variable of a GPU fail time, then its
probability density function is $f(t)$ and its cumulative distribution
function $F(t) = Pr\{T < t\}$ gives the probability that a GPU fails
by duration $t$. The survival function is then 
\begin{displaymath}
  S(t) = Pr\{T \geq t\} = 1 - F(t) = \int_t^\infty f(x)dx,
\end{displaymath}
which is the probability of being operational at duration $t$.  We use
the R packages \pkg{survival} and \pkg{survminer} for the KM analysis,
which is reported in Figure~\ref{fig:km-all-cage-node}. Within each
batch, separate survival curves are computed for each cage by node
combination.

Another standard survival data methodology is a Cox proportional
hazards (CPH) regression analys, which can include covariates and
estimate relative risk based on the covariates
\cite{Cox1972,Harrell2015}. The CPH regression function takes the form
\begin{displaymath}
  H_k(t) = H_0(t)e^{\sum\limits_{i=1}^n{\beta x}}
\end{displaymath}
That is, a base hazard rate multiplied by a function of
covariates. The key assumption in the CPH model is that the "hazards"
are multipliers on the "baseline hazard" but the estimate of the
baseline hazard is nonparametric, meaning it makes no specific
distributional assumption and just learns from the data (unlike
analyses that assume for example a Weibul or an exponential
model). The multiplicative assumption is that the hazard curves do not
cross and are multiples of each other. The multiplier is the *hazard
coefficient*. For example, if the baseline is *node0* and the *node2*
hazard is 2, then *node2* sees twice as many failures as *node0* on
average.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\columnwidth]{figs/cox_o001.pdf}
  \caption{GPU hazard ratios from Cox regression model on {\tt old}
    batch.}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=0.8\columnwidth]{figs/cox_n001.pdf}
  \caption{GPU hazard ratios from Cox regression model on {\tt new}
    batch.}
\end{figure}
